---
title: "Course Evaluation Comparison Dashboard"
format: 
  dashboard:
    orientation: columns
    theme: cosmo
    scrolling: true
server: shiny
---

```{r}
#| context: setup
library(shiny)
library(ggplot2)
library(dplyr)
library(tidyr)
library(plotly)
library(DT)

# Load all course data
source(file="read_data_save.R")
```

# Overview {data-icon="fa-chart-line"}

## Column {width=250}

### Filters

```{r}
selectInput("selected_courses", "Select Courses to Compare:",
            choices = names(courses),
            selected = names(courses)[1],
            multiple = TRUE)

selectInput("class_standing_filter", "Class Standing:",
            choices = c("All", "Freshman/first year", "Sophomore/second year", 
                       "Junior/third year", "Senior/ fourth year", "Graduate"),
            selected = "All")

selectInput("attendance_filter", "Attendance:",
            choices = c("All", "75-100%", "50-74%", "25-49%", "0-24%"),
            selected = "All")
```

### Overall Summary

```{r}
#| title: "Course Summary"
tableOutput("overall_summary")
```

## Column {width=550}

### Response Summary by Course

```{r}
DTOutput("response_summary_table")
```

### Class Standing Comparison

```{r}
plotlyOutput("class_standing_comparison", height = "300px")
```

### Attendance Comparison

```{r}
plotlyOutput("attendance_comparison", height = "300px")
```

# Ratings {data-icon="fa-star"}

## Column

### Comparative Rating Profiles

```{r}
plotlyOutput("comparative_radar", height = "600px")
```

### Average Ratings by Course

```{r}
plotlyOutput("ratings_comparison_bar", height = "400px")
```

# Crosstabs {data-icon="fa-table"}

## Column {width=230}

### Analysis Options

```{r}
selectInput("crosstab_courses", "Select Courses:",
            choices = names(courses),
            selected = names(courses)[1],
            multiple = TRUE)

selectInput("rating_var", "Select Rating Variable:",
            choices = c("Time Effective" = "time_effective",
                       "Explained Concepts" = "explained_concepts",
                       "Useful Feedback" = "useful_feedback",
                       "Grading Clear" = "grading_clear",
                       "Felt Engaged" = "felt_engaged",
                       "Lectures Structured" = "lectures_structured",
                       "Activities Structured" = "activities_structured",
                       "Homework Helpful" = "homework_helpful"))

selectInput("student_var", "Select Student Variable:",
            choices = c("Class Standing" = "class_standing",
                       "Attendance" = "attendance_pct",
                       "Hours per Week" = "hours_per_week",
                       "Why Taking Course" = "why_taking",
                       "Felt Prepared" = "felt_prepared"))
```

## Column {width=670}

### Crosstabulation Table

```{r}
DTOutput("crosstab_table")
```

### Visualization by Course

```{r}
plotlyOutput("crosstab_plot", height = "400px")
```

### Summary Statistics

```{r}
tableOutput("crosstab_summary")
```

```{r}
#| context: server

# Helper function to convert Likert scale to numeric
likert_to_numeric <- function(x) {
  case_when(
    x == "Very frequently" ~ 5,
    x == "Frequently" ~ 4,
    x == "Somewhat frequently" ~ 3,
    x == "Occasionally" ~ 2,
    x == "Rarely" ~ 1,
    x == "Unable to comment" ~ NA_real_,
    TRUE ~ NA_real_
  )
}

# Reactive filtered data for overview
filtered_overview_data <- reactive({
  req(input$selected_courses)
  
  # Combine selected courses with course identifier
  combined <- bind_rows(
    lapply(input$selected_courses, function(course_name) {
      df <- courses[[course_name]]
      df$course_name <- course_name
      return(df)
    })
  )
  
  # Apply filters
  if(input$class_standing_filter != "All") {
    combined <- combined %>% filter(class_standing == input$class_standing_filter)
  }
  
  if(input$attendance_filter != "All") {
    combined <- combined %>% filter(attendance_pct == input$attendance_filter)
  }
  
  return(combined)
})

# Overall summary
output$overall_summary <- renderTable({
  df <- filtered_overview_data()
  
  n_courses <- length(unique(df$course_name))
  n_students <- nrow(df)
  
  # Determine if graduate or undergraduate
  grad_count <- sum(df$class_standing == "Graduate", na.rm = TRUE)
  undergrad_count <- n_students - grad_count
  
  data.frame(
    Metric = c("Total Courses", "Total Students", "Undergraduate", "Graduate"),
    Value = c(n_courses, n_students, undergrad_count, grad_count)
  )
})

# Response summary by course
output$response_summary_table <- renderDT({
  df <- filtered_overview_data()
  
  summary_data <- df %>%
    group_by(course_name) %>%
    summarise(
      `Total Responses` = n(),
      `Most Common Attendance` = names(which.max(table(attendance_pct))),
      `Most Common Hours/Week` = names(which.max(table(hours_per_week))),
      `Avg Rating` = round(mean(c(
        mean(likert_to_numeric(time_effective), na.rm = TRUE),
        mean(likert_to_numeric(explained_concepts), na.rm = TRUE),
        mean(likert_to_numeric(felt_engaged), na.rm = TRUE)
      ), na.rm = TRUE), 1),
      .groups = 'drop'
    )
  
  summary_data
}, options = list(pageLength = 10, scrollX = TRUE, dom = 't'))

# Class standing comparison
output$class_standing_comparison <- renderPlotly({
  df <- filtered_overview_data() %>%
    filter(!is.na(class_standing)) %>%
    count(course_name, class_standing)
  
  plot_ly(df, x = ~class_standing, y = ~n, color = ~course_name, 
          type = 'bar') %>%
    layout(title = list(text = "Class Standing Distribution", font = list(size = 14)),
           xaxis = list(title = "Class Standing", tickangle = -45),
           yaxis = list(title = "Number of Students"),
           barmode = 'group',
           margin = list(l = 50, r = 20, t = 50, b = 120),
           legend = list(orientation = 'h', y = -0.3))
})

# Attendance comparison
output$attendance_comparison <- renderPlotly({
  df <- filtered_overview_data() %>%
    filter(!is.na(attendance_pct)) %>%
    count(course_name, attendance_pct)
  
  plot_ly(df, x = ~attendance_pct, y = ~n, color = ~course_name,
          type = 'bar') %>%
    layout(title = list(text = "Attendance Distribution", font = list(size = 14)),
           xaxis = list(title = "Attendance (%)"),
           yaxis = list(title = "Count"),
           barmode = 'group',
           margin = list(l = 50, r = 20, t = 50, b = 100),
           legend = list(orientation = 'h', y = -0.25))
})

# Comparative radar plot
output$comparative_radar <- renderPlotly({
  req(input$selected_courses)
  
  df <- filtered_overview_data()
  
  # Calculate ratings for each course
  ratings_by_course <- df %>%
    group_by(course_name) %>%
    summarise(
      time_use = mean(likert_to_numeric(time_effective), na.rm = TRUE),
      explanation = mean(likert_to_numeric(explained_concepts), na.rm = TRUE),
      feedback = mean(likert_to_numeric(useful_feedback), na.rm = TRUE),
      grading = mean(likert_to_numeric(grading_clear), na.rm = TRUE),
      engagement = mean(likert_to_numeric(felt_engaged), na.rm = TRUE),
      structure = mean(likert_to_numeric(lectures_structured), na.rm = TRUE),
      materials = mean(likert_to_numeric(homework_helpful), na.rm = TRUE),
      .groups = 'drop'
    )
  
  categories <- c("Time Use", "Explanation", "Feedback", "Grading", 
                  "Engagement", "Structure", "Materials")
  
  # Create radar plot with multiple courses
  fig <- plot_ly(type = 'scatterpolar', fill = 'toself', mode = 'lines+markers')
  
  colors <- c('steelblue', 'coral', 'mediumseagreen', 'mediumpurple', 'gold', 'lightcoral')
  
  for(i in 1:nrow(ratings_by_course)) {
    values <- as.numeric(ratings_by_course[i, -1])
    fig <- fig %>%
      add_trace(
        r = values,
        theta = categories,
        name = ratings_by_course$course_name[i],
        marker = list(color = colors[i %% length(colors) + 1]),
        line = list(color = colors[i %% length(colors) + 1])
      )
  }
  
  fig <- fig %>%
    layout(
      polar = list(
        radialaxis = list(
          visible = TRUE,
          range = c(0, 5)
        )
      ),
      title = list(text = "Course Rating Profiles Comparison", font = list(size = 16)),
      margin = list(l = 80, r = 80, t = 80, b = 80),
      legend = list(orientation = 'h', y = -0.15)
    )
  
  fig
})

# Ratings comparison bar chart
output$ratings_comparison_bar <- renderPlotly({
  df <- filtered_overview_data()
  
  ratings_long <- df %>%
    mutate(across(c(time_effective, explained_concepts, useful_feedback, 
                    grading_clear, felt_engaged, lectures_structured, 
                    homework_helpful), likert_to_numeric)) %>%
    group_by(course_name) %>%
    summarise(
      `Time Use` = mean(time_effective, na.rm = TRUE),
      `Explanation` = mean(explained_concepts, na.rm = TRUE),
      `Feedback` = mean(useful_feedback, na.rm = TRUE),
      `Grading` = mean(grading_clear, na.rm = TRUE),
      `Engagement` = mean(felt_engaged, na.rm = TRUE),
      `Structure` = mean(lectures_structured, na.rm = TRUE),
      `Materials` = mean(homework_helpful, na.rm = TRUE),
      .groups = 'drop'
    ) %>%
    pivot_longer(-course_name, names_to = "Category", values_to = "Rating")
  
  plot_ly(ratings_long, x = ~Category, y = ~Rating, color = ~course_name,
          type = 'bar') %>%
    layout(title = list(text = "Average Ratings by Category", font = list(size = 16)),
           xaxis = list(title = "", tickangle = -30),
           yaxis = list(title = "Average Rating (1-5)", range = c(0, 5)),
           barmode = 'group',
           margin = list(l = 60, r = 40, t = 60, b = 120),
           legend = list(orientation = 'h', y = -0.3))
})

# Reactive filtered data for crosstabs
filtered_crosstab_data <- reactive({
  req(input$crosstab_courses)
  
  # Combine selected courses
  combined <- bind_rows(
    lapply(input$crosstab_courses, function(course_name) {
      df <- courses[[course_name]]
      df$course_name <- course_name
      return(df)
    })
  )
  
  return(combined)
})

# Crosstabulation table
output$crosstab_table <- renderDT({
  df <- filtered_crosstab_data() %>%
    mutate(rating_numeric = likert_to_numeric(.data[[input$rating_var]])) %>%
    select(course_name, student_var = all_of(input$student_var), 
           rating_var = all_of(input$rating_var),
           rating_numeric)
  
  # Create frequency table by course
  freq_table <- df %>%
    count(course_name, student_var, rating_var) %>%
    pivot_wider(names_from = rating_var, values_from = n, values_fill = 0) %>%
    arrange(course_name, student_var)
  
  freq_table
}, options = list(pageLength = 15, scrollX = TRUE, autoWidth = TRUE))

# Crosstab visualization
output$crosstab_plot <- renderPlotly({
  df <- filtered_crosstab_data() %>%
    mutate(rating_numeric = likert_to_numeric(.data[[input$rating_var]])) %>%
    filter(!is.na(rating_numeric)) %>%
    group_by(course_name, student_var = .data[[input$student_var]]) %>%
    summarise(
      mean_rating = mean(rating_numeric, na.rm = TRUE),
      count = n(),
      .groups = 'drop'
    ) %>%
    filter(!is.na(student_var))
  
  plot_ly(df, x = ~student_var, y = ~mean_rating, color = ~course_name,
          type = 'bar',
          text = ~paste("Mean:", round(mean_rating, 2), "<br>N:", count),
          hoverinfo = 'text') %>%
    layout(
      title = list(text = paste("Average", input$rating_var, "by", input$student_var), 
                   font = list(size = 14)),
      xaxis = list(title = input$student_var, tickangle = -30),
      yaxis = list(title = "Average Rating", range = c(0, 5)),
      barmode = 'group',
      margin = list(l = 60, r = 40, t = 60, b = 120),
      legend = list(orientation = 'h', y = -0.3)
    )
})

# Crosstab summary statistics
output$crosstab_summary <- renderTable({
  df <- filtered_crosstab_data() %>%
    mutate(rating_numeric = likert_to_numeric(.data[[input$rating_var]])) %>%
    filter(!is.na(rating_numeric)) %>%
    group_by(course_name, student_var = .data[[input$student_var]]) %>%
    summarise(
      N = n(),
      Mean = round(mean(rating_numeric, na.rm = TRUE), 2),
      SD = round(sd(rating_numeric, na.rm = TRUE), 2),
      Min = min(rating_numeric, na.rm = TRUE),
      Max = max(rating_numeric, na.rm = TRUE),
      .groups = 'drop'
    ) %>%
    filter(!is.na(student_var))
  
  df
}, striped = TRUE, hover = TRUE)
```
